An Intrinsically-Typed Solution for the\\ List-Machine Benchmark

### 2021-03-09 papers/COLA/paper.tex
Fixes on the code

In this section we describe the necessary modifications to our formalization to include indirect jumps
as proposed by Appel et. al.~\cite{AppelDL12}. We start by describing the changes on the type system
and its impact on the intrinsically typed representation of programs in Section~\ref{sec:changestypes}.
Updates on the type checker and interpreter implementations are described in Section~\ref{sec:changesintep}.

\subsection{Modifications in the type system and instruction set}

@@

\subsection{Modifications in the type-checker and interpreter}\label{sec:changesintep}

>>>>

In this section we describe the necessary modifications to our formalization to include indirect jumps
as proposed by Appel et. al.~\cite{AppelDL12}. We start by describing the changes on the type system
and its impact on the intrinsically typed representation of programs in Section~\ref{sec:changestypes}.
Updates on the type checker and interpreter implementations are described in Sections~\ref{sec:changestypechecker}
and ~\ref{sec:changesintep}, respectively.

\subsection{Modifications in the type system and instruction set}\label{sec:changestypes}

@@

Once we have modified the subtyping and least upper bound relations, most of the type system
rules remain unchanged from the original version. In order to support indirect jumps, Appel et. al.
approach is to modify the jump instruction and add a new instruction for loading label values into
machine registers. The typing rules for get-label and jump instructions are as follows.

\[
\begin{array}{cc}
  \infer{\Pi\vdash_{instr} \Gamma \{\textrm{get-label}\:v\:L_0\}\Gamma'}
        {\Gamma[v\,:=\,nil] = \Gamma'} &
  \infer{\Pi\vdash_{instr} \Gamma \{\textrm{get-label}\:v\:l\}\Gamma'}
        {\Pi(l) = \Gamma_1 & \Gamma[v\,:=\,cont\:\Gamma_1] = \Gamma'} \\ & \\
  \multicolumn{2}{c}{
    \infer{\Pi;\Gamma\vdash_{block}\,\textrm{jump}\:v}
          {\Gamma(v) = cont\:\Gamma_1 & \Gamma \subset \Gamma_1}
  }
\end{array}
\]
The first rule show that the type of label $L_0$, the starting label of a complete machine program,
is $nil$ and the second specifies that the type for any other label is a continuation type formed by
the context associated with label $l$ in $\Pi$. The last rule show that typing a jump instruction
requires that its register has a continuation type, $cont\:\Gamma_1$, that should be a subtype of
current block context $\Gamma$. Based on these rules, we add the following constructors to our
instruction typed syntax:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
\column{35}{@{}>{\hspre}l<{\hspost}@{}}%
\column{40}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathkw{data}\;\D{\_\vdash\_\Rightarrow\_}\;(\V{\Pi}\;\mathbin{:}\;\D{PCtx})\;(\V{\Gamma}\;\mathbin{:}\;\D{Ctx})\;\mathbin{:}\;\D{Ctx}\;\rightarrow\;\D{Set}\;\mathkw{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\orange{\texttt{-- unchanged from the previous code version.}}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Con{instr\textrm{-}getlabel\textrm{-}0}\;{}\<[25]%
\>[25]{}\mathbin{:}\;\D{\forall}\;\{\mskip1.5mu \V{\tau}\;\V{x}\mskip1.5mu\}\;\rightarrow\;(\V{idx}\;\mathbin{:}\;(\V{x}\;\red{,}\,\;\V{\tau})\;\D{\in}\;\V{\Gamma})\;{}\<[E]%
\\
\>[25]{}\hsindent{10}{}\<[35]%
\>[35]{}\rightarrow\;\V{\Pi}\;\D{\vdash}\;\V{\Gamma}\;\D{\Rightarrow}\;(\V{idx}\;\F{::=}\;(\V{x}\;\red{,}\,\;\Con{nil})){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Con{instr\textrm{-}getlabel}\;{}\<[25]%
\>[25]{}\mathbin{:}\;\D{\forall}\;\{\mskip1.5mu \V{l}\;\V{\tau}\;\V{\Gamma_1}\;\V{x}\mskip1.5mu\}\;\rightarrow\;(\V{idx}\;\mathbin{:}\;(\V{x}\;\red{,}\,\;\V{\tau})\;\D{\in}\;\V{\Gamma})\;{}\<[E]%
\\
\>[25]{}\hsindent{15}{}\<[40]%
\>[40]{}\rightarrow\;\V{\Pi}\;\D{[}\;\V{l}\;\D{]=}\;\V{\Gamma_1}\;{}\<[E]%
\\
\>[25]{}\hsindent{15}{}\<[40]%
\>[40]{}\rightarrow\;\V{\Pi}\;\D{\vdash}\;\V{\Gamma}\;\D{\Rightarrow}\;(\V{idx}\;\F{::=}\;(\V{x}\;\red{,}\,\;\Con{cont}\;\V{\Gamma_1})){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
Constructor \ensuremath{\Con{instr\textrm{-}getlabel\textrm{-}0}} encodes the typing rule for the instruction \textrm{get-label} when
the label involved is the special label $L_0$. Intutively, the constructor \ensuremath{\Con{instr\textrm{-}getlabel\textrm{-}0}} type
ensures that the type associated with the register for the instruction is the empty list type.
Next, constructor \ensuremath{\Con{instr\textrm{-}getlabel}} show that any other label $l$ should be typed with a continuation
holding the typing context associated with $l$ in $\Pi$.

The needed modifications on block syntax are presented next.
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
\column{42}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathkw{data}\;\D{Block}\;(\V{\Pi}\;\mathbin{:}\;\D{PCtx})\;(\V{\Gamma}\;\mathbin{:}\;\D{Ctx})\;\mathbin{:}\;\D{Ctx}\;\rightarrow\;{}\<[42]%
\>[42]{}\D{Set}\;\mathkw{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\orange{\texttt{-- unchanged from the previous code version}}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Con{block\textrm{-}jump}\;{}\<[25]%
\>[25]{}\mathbin{:}\;\D{\forall}\;\{\mskip1.5mu \V{l}\;\V{x}\;\V{\Gamma_1}\mskip1.5mu\}\;\rightarrow\;(\V{x}\;\red{,}\,\;\Con{cont}\;\V{\Gamma_1})\;\D{\in}\;\V{\Gamma}\;{}\<[E]%
\\
\>[25]{}\rightarrow\;\V{\Gamma}\;\D{\subset}\;\V{\Gamma_1}\;{}\<[E]%
\\
\>[25]{}\rightarrow\;\D{Block}\;\V{\Pi}\;\V{\Gamma}\;\V{\Gamma_1}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
We only need to adapt the type of constructor for jump by requiring that the register has a
continuation type whose typing context is a subtype of the current block type, which allows
the safe jump to the next block.

\subsection{Modifications in the type-checker}\label{sec:changestypechecker}

\subsection{Modification in the intepreter}\label{sec:changesintep}
